Tailoring Differentially Private Bayesian Inference to Distance Between Distributions

### 2018-10-21 Arxiv/main.tex
rename folder
update arvix plots and texts

\section{Motivation}
\label{sec_intro}
Publishing the posterior distribution inferred from a sensitive dataset can
leak information about the individuals in the dataset.
In order to guarantee differential privacy and to protect the
individuals' data we can add noise to the posterior before releasing it.
The amount of the noise that we need to  introduced
depends on the privacy parameter $\epsilon$ and the sensitivity of the inference to
small changes in the data set. 
Sensitivity can be computed in many different ways based on which metric space
we consider on the output set of the mechanism. In the literature on private Bayesian
inference (\cite{zhang2016differential,xiao2012bayesian}), it is only measured with
respect to the vector of numbers parametrizing the output distribution using, e.g. the $\ell_1$ norm.
A more natural approach which we explore here, is to measure sensitivity with respect to a metric on the space of inferred probability distributions.
A re-loved question is that of how to measure accuracy. Again,
this can be answered in different ways based on the metric imposed on the output space, and yet again
only in few works in literature (e.g. \cite{zhang2016differential})
distances between probability measures have been used for these purposes.

@@

\begin{definition}
\label{def_smoo}
The mechanism $\hexpmech(x)$ outputs a candidate $r \in \candidateset$ with probability
\begin{equation*}
\underset{z \thicksim \hexpmech}{\Pr}[z=r] = \frac {exp\big(\frac{-\epsilon\cdot\hellinger(\bysinfer(\dataobs),r)}{2\cdot S(\dataobs)}\big)}
{\displaystyle\sum_{r\in\candidateset} exp\Big(\frac{-\epsilon\cdot\hellinger(\bysinfer(\dataobs),r)}{2\cdot S(\dataobs)}\Big)}.
\end{equation*}
where $S_\beta(x)$ is the smooth sensitivity of $\hlg(\bysinfer(x),-)$, calculated by:

@@

In what follows, we will use a correspondence between the probability
 $\underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$ of every
 $r\in\candidateset$ and the probability 
 $\underset{z \thicksim \hexpmech(x)}{Pr}[\hlg(\bysinfer(x),z) =
 \hlg(\bysinfer(x),r)]$ for the utility score for $r$. In particular, for every
 $r\in\candidateset$ we have:

$$
\underset{z \thicksim \hexpmech(x)}{Pr}[z = r]=
\frac{1}{2}\Big (\underset{z \thicksim \hexpmech(x)}{Pr}[\hlg(\bysinfer(x),z) =
 \hlg(\bysinfer(x),r)]\Big )
$$

To see this, it is enough to notice that: $\underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$ is proportional too $\hlg(\bysinfer(x),r)$, i.e., $u(x,z)$. We can derive, if $u(r,x) = u(r',x)$ then $\underset{z \thicksim \hexpmech(x)}{Pr}[z = r] = \underset{z \thicksim \hexpmech(x)}{Pr}[z = r']$. We assume the number of candidates $z \in \mathcal{R}$ that satisfy $u(z,x) = u(r,x)$ is $|r|$, we have  $\underset{z \thicksim \hexpmech(x)}{Pr}[u(z,x) = u(r,x)] = |r| \underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$. Because Hellinger distance  $\hlg(\bysinfer(x),z)$ is axial symmetry, where the $\bysinfer(x)$ is the symmetry axis. It can be infer that $|z| = 2$ for any candidates, apart from the true output, i.e., $\underset{z \thicksim \hexpmech(x)}{Pr}[u(z,x) = u(r,x)] = 2 \underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$. This parameter can be eliminate in both sides in proof.

@@

The sensitivity is always greater than 0, and our utility function $-\hlg(\bysinfer(x),z)$ is smaller than zero, i.e., $u(z,x) \leq 0$, we need to consider two cases where $\lambda < 0$, and $\lambda > 0$:

@@

\section{Accuracy Analysis}
\subsection{Laplace Mechanism}
\label{subsec_accuracy_lap}
Fixing a data set $x$, we will sample noise $Lap_i = floor(Y)$ where $Y \sim Lap(\frac{2}{\epsilon})$. Since already had the accuracy bound based on the $l_1$ norm in Laplace mechanism:
\begin{equation*}
Pr[|Y| \geq t] = e^{- \frac{t \epsilon}{2}}.
\end{equation*}

So we have the probability $Pr[| Lap_i |] = Pr[| Lap_i | \leq |Y| < | Lap_i | + 1] = e^{- \frac{| Lap_i | \epsilon}{2}} - e^{- \frac{(| Lap_i | + 1) \epsilon}{2}}$.

When we are in the case of $m$ dimension Dirichlet distribution, we will add $(m-1)$ i.i.d. Laplace noises $\{| Lap_1 |, | Lap_2 |, \cdots, | Lap_{m-1} |\}$ to the output. Then the probability of each group of noises are calculated as $Pr[\{| Lap_1 |, | Lap_2 |, \cdots, | Lap_{m-1} |\}] = Pr[| Lap_1 | \leq |Y| < | Lap_1 | + 1] \times Pr[| Lap_2 | \leq |Y| < | Lap_2 | + 1] \times \cdots \times Pr[| Lap_{m-1} | \leq |Y| < | Lap_{m-1} | + 1] = (e^{- \frac{| Lap_1 | \epsilon}{2}} - e^{- \frac{(| Lap_1 | + 1) \epsilon}{2}}) \times (e^{- \frac{| Lap_2 | \epsilon}{2}} - e^{- \frac{(| Lap_2 | + 1) \epsilon}{2}}) \times \cdots \times (e^{- \frac{| Lap_{m-1} | \epsilon}{2}} - e^{- \frac{(| Lap_{m-1} | + 1) \epsilon}{2}})$.

@@

First is the tight bound with very accurate calculation.
\begin{equation*}
\underset{z \thicksim \hexpmech(x)}{Pr}[ \hlg(\bysinfer(x), z) \geq c] = \sum\limits_{\{z | \hlg(\bysinfer(x), z) \geq c\}} \frac{e^{\frac{- \epsilon \hlg(\bysinfer(x), z)}{S(x)}}}{NL_x}.
\end{equation*}

@@

\begin{equation*}
\underset{z \thicksim \hexpmech(x)}{Pr}[ \hlg(\bysinfer(x), z) \geq c] \leq \frac{|R| \exp{(\frac{- \epsilon c}{S(x)}})}{NL_x}.
\end{equation*}

@@

\begin{equation*}
\underset{z \thicksim \hexpmech(x)}{Pr}[ \hlg(\bysinfer(x), z) \geq c] \leq \ \frac{|R| \exp{(\frac{- \epsilon c}{S(x)}})}{N(n)},
\end{equation*}
where we replace the $NL_x$ with a value only related to the size of the data. However, we haven't figured out the formula of this $N(n)$.

@@

\subsection{Computation Efficiency}
\label{subsec_effi}
The formula for computing the local sensitivity presented in Sec. \ref{subsec_emls}:$\max\limits_{\{|x,y'| \leq 1; y'\in \mathcal{X}^n\}}\max\limits_{\{r\in R\}} \{\hlg(\bysinfer(x), r) - \hlg(\bysinfer(y'), r)|\}$ can be reduced to $\max\limits_{\{|x,y'| \leq 1;y'\in \mathcal{X}^n\}}\hlg(\bysinfer(x), \bysinfer(y'))|$ by applying the distance triangle property. i.e., the maximum value over $\max\limits_{r \in R}$ always happen when $r = \bysinfer(x)$ itself, where $\Delta_{l}u(x) = \max\limits_{\{|x,y'| \leq 1;y'\in \mathcal{X}^n\}} \{\hlg(\bysinfer(x), \bysinfer(x)) - \hlg(\bysinfer(y'), \bysinfer(x))|\} = \max\limits_{\{|x,y'| \leq 1;y'\in \mathcal{X}^n\}} \{\hlg(\bysinfer(y'), \bysinfer(x))|\}$. We also have some experiments for validating our proposal as in Fig. \ref{fig_efficiency}, where we calculate the $\max\limits_{\{|x,y'| \leq 1;y'\in \mathcal{X}^n\}}$ value for every candidate $r \in R$. It is shown that maximum value taken when  $r = \bysinfer(x)$.

@@

\subsection{Accuracy Trade-off Evaluation wrt. Different Variables}
\label{subsec_vs_variables}
In this section, we evaluate the accuracy of the mechanisms defined in
Section (\ref{sec_smoo}) w.r.t. four variables, including data size, dimensions,
data variance, prior distribution, and some combinations thereof.
Every plot is an average over 1000 runs. In all the experiments we set
$\epsilon = 0.8$, and $\delta = 10^{-8}$.

@@

\begin{figure}[H]
\begin{center}
\centering
  \subfigure[Data set size from $300$ to $800$ ]{
    \includegraphics[width=0.45\textwidth]{accuracy_vs_datasize_800.eps}
    \label{subsubsec_vs_datasize1a}
  }
  \subfigure[Data set size from $14000$ to $20000$]{
    \includegraphics[width=0.45\textwidth]{accuracy_vs_datasize_20000.eps}
  \label{subsubsec_vs_datasize1b}} 
\caption{Increasing data size with fixed prior $\betad(1,1)$. Unbalanced datasets of mean $(0.1,0.9)$ and parameters $\epsilon = 0.8$ and $\delta = 10^{-8}$}
\label{fig_vs_datasize}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\centering
  % \subfigure[Data set size from $300$ to $800$ ]{
  %   \includegraphics[width=0.45\textwidth]{accuracy_vs_datasize_300_dirichlet.eps}
  %   \label{subsubsec_vs_datasize1adir}
  % }
  \subfigure[Data set size from $14000$ to $20000$]{
     \includegraphics[width=0.45\textwidth]{accuracy_vs_datasize_1k_dirichlet.eps}   
  \label{subsubsec_vs_datasize1bdir}
  } 
\caption{Increasing data size with fixed $\dirichlet(1,1,1)$ prior distribution, Unbalanced datasets of mean $(0.2,0.3, 0.5)$ and parameters $\epsilon = 0.8$ and $\delta = 10^{-8}$}
\label{fig_vs_datasize_dir}
\end{center}
\end{figure}

@@

\paragraph{Fixed dataset varying balanced priors}
\label{subsubsec_vs_prior}
In Figure \ref{fig_vs_prior}, we fix the dataset to be $\langle 5,5,5\rangle$.
We also considered balanced priors with increasing values in their dimensions.
The plot shows that in the beginning the Laplace mechanism performs better but
it is outperformed after a while.
\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{accuracy_vs_prior_5_5_5.eps}
\caption{Observed data set is: $(5,5,5)$, varying balanced priors}
\label{fig_vs_prior}
\end{figure}

@@

\begin{figure*}
\begin{center}
\centering
  \subfigure[data size range from 90 to 180]{
    \includegraphics[width=0.45\textwidth]{global_epsilon_20_20.eps}}
  \subfigure[data size range from 90 to 180]{
    \includegraphics[width=0.45\textwidth]{global_epsilon_20_20(2).eps}} 
\caption{Concrete privacy calculation under settings that: prior distribution:$[1,1]$, $\epsilon = 0.8$, $\delta = 0.0005$ and observed data are uniformly distributed}
\label{fig_privacy}
\end{center}
\end{figure*}

>>>>

\section{Motivation}
\label{sec_moti}
Publishing the posterior distribution inferred from a sensitive dataset can
leak information about the individuals in the dataset.
In order to guarantee differential privacy and to protect the
individuals' data we can add noise to the posterior before releasing it.
The amount of the noise that we need to  introduced
depends on the privacy parameter $\epsilon$ and the sensitivity of the inference to
small changes in the data set. 
Sensitivity can be computed in many different ways based on which metric space
we consider on the output set of the mechanism. In the literature on private Bayesian
inference (\cite{zhang2016differential,xiao2012bayesian}), it is only measured with
respect to the vector of numbers parametrizing the output distribution using, e.g. the $\ell_1$ norm.
A more natural approach which we explore here, is to measure sensitivity with respect to a metric on the space of inferred probability distributions.
A re-loved question is that of how to measure accuracy. Again,
this can be answered in different ways based on the metric imposed on the output space, and yet again
only in few works in literature (e.g. \cite{zhang2016differential})
distances between probability measures have been used for these purposes.

@@

\begin{definition}
\label{def_smoo}
The mechanism $\hexpmech(x)$ outputs a candidate $r \in \candidateset$ with probability
\begin{equation*}
\underset{z \thicksim \hexpmech}{\Pr}[z=r] = \frac {exp\big(\frac{-\epsilon\cdot\hellinger(\bysinfer(\dataobs),r)}{2\cdot S(\dataobs)}\big)}
{\displaystyle\sum_{r\in\candidateset} exp\Big(\frac{-\epsilon\cdot\hellinger(\bysinfer(\dataobs),r)}{2\cdot S(\dataobs)}\Big)}.
\end{equation*}
where $S_\beta(x)$ is the smooth sensitivity of $\hellinger(\bysinfer(x),-)$, calculated by:

@@

In what follows, we will use a correspondence between the probability
 $\underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$ of every
 $r\in\candidateset$ and the probability 
 $\underset{z \thicksim \hexpmech(x)}{Pr}[\hellinger(\bysinfer(x),z) =
 \hellinger(\bysinfer(x),r)]$ for the utility score for $r$. In particular, for every
 $r\in\candidateset$ we have:

$$
\underset{z \thicksim \hexpmech(x)}{Pr}[z = r]=
\frac{1}{2}\Big (\underset{z \thicksim \hexpmech(x)}{Pr}[\hellinger(\bysinfer(x),z) =
 \hellinger(\bysinfer(x),r)]\Big )
$$

To see this, it is enough to notice that: $\underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$ is proportional too $\hellinger(\bysinfer(x),r)$, i.e., $u(x,z)$. We can derive, if $u(r,x) = u(r',x)$ then $\underset{z \thicksim \hexpmech(x)}{Pr}[z = r] = \underset{z \thicksim \hexpmech(x)}{Pr}[z = r']$. We assume the number of candidates $z \in \mathcal{R}$ that satisfy $u(z,x) = u(r,x)$ is $|r|$, we have  $\underset{z \thicksim \hexpmech(x)}{Pr}[u(z,x) = u(r,x)] = |r| \underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$. Because Hellinger distance  $\hellinger(\bysinfer(x),z)$ is axial symmetry, where the $\bysinfer(x)$ is the symmetry axis. It can be infer that $|z| = 2$ for any candidates, apart from the true output, i.e., $\underset{z \thicksim \hexpmech(x)}{Pr}[u(z,x) = u(r,x)] = 2 \underset{z \thicksim \hexpmech(x)}{Pr}[z = r]$. This parameter can be eliminate in both sides in proof.

@@

The sensitivity is always greater than 0, and our utility function $-\hellinger(\bysinfer(x),z)$ is smaller than zero, i.e., $u(z,x) \leq 0$, we need to consider two cases where $\lambda < 0$, and $\lambda > 0$:

@@

\section{Accuracy Analysis}

\subsection{Accuracy Bound for Laplace Mechanism}
\label{subsec_accuracy_lap}
Accuracy bound for Laplace mechanism is provided by its probability density function:
\begin{equation*}
Pr[|Y| \geq t] = e^{- \frac{t}{b}},
\end{equation*}
where $Y \sim Lap(b)$, $b = \frac{GS}{\epsilon}$

After post-processing:  $Pr[\lfloor Y \rfloor = t] = Pr[ t-1 \leq Y < t] = \frac{1}{2} (e^{- \frac{t - 1}{b}} - e^{- \frac{t}{b}})$.

@@

First is the tight bound with very accurate calculation.
\begin{equation*}
\underset{z \thicksim \hexpmech(x)}{Pr}[ \hellinger(\bysinfer(x), z) \geq c] = \sum\limits_{\{z | \hellinger(\bysinfer(x), z) \geq c\}} \frac{e^{\frac{- \epsilon \hellinger(\bysinfer(x), z)}{S(x)}}}{NL_x}.
\end{equation*}

@@

\begin{equation*}
\underset{z \thicksim \hexpmech(x)}{Pr}[ \hellinger(\bysinfer(x), z) \geq c] \leq \frac{|R| \exp{(\frac{- \epsilon c}{S(x)}})}{NL_x}.
\end{equation*}

@@

\begin{equation*}
\underset{z \thicksim \hexpmech(x)}{Pr}[ \hellinger(\bysinfer(x), z) \geq c] \leq \ \frac{|R| \exp{(\frac{- \epsilon c}{S(x)}})}{N(n)},
\end{equation*}
where we replace the $NL_x$ with a value only related to the size of the data. However, we haven't figured out the formula of this $N(n)$.

@@

\subsection{Computation Efficiency}
\label{subsec_effi}
The formula for computing the local sensitivity presented in Sec. \ref{subsec_emls}:$\max\limits_{\{|x,y'| \leq 1; y'\in \mathcal{X}^n\}}\max\limits_{\{r\in R\}} \{\hellinger(\bysinfer(x), r) - \hellinger(\bysinfer(y'), r)|\}$ can be reduced to $\max\limits_{\{|x,y'| \leq 1;y'\in \mathcal{X}^n\}}\hellinger(\bysinfer(x), \bysinfer(y'))|$ by applying the distance triangle property. i.e., the maximum value over $\max\limits_{r \in R}$ always achieves at $r = \bysinfer(x)$ itself, where $\Delta_{l}u(x) = \max\limits_{\{|x,y'| \leq 1;y'\in \mathcal{X}^n\}} \{\hellinger(\bysinfer(x), \bysinfer(x)) - \hellinger(\bysinfer(y'), \bysinfer(x))|\} = \max\limits_{\{|x,y'| \leq 1;y'\in \mathcal{X}^n\}} \{\hellinger(\bysinfer(y'), \bysinfer(x))|\}$. We also have some experiments for validating our proposal as in Fig. \ref{fig_efficiency}, where we calculate the $\max\limits_{\{|x,y'| \leq 1;y'\in \mathcal{X}^n\}}$ value for every candidate $r \in R$. It is shown that maximum value taken when  $r = \bysinfer(x)$.

@@

\subsection{Theoretical Results}

\begin{figure}[H]
\begin{center}
\centering
  \subfigure[2 dimensions with data size $600$]{
    \includegraphics[width=0.22\textwidth]{theory_2d.eps}
  \label{subfig_theory_2d}
  }
  \subfigure[3 dimensions with data size $600$]{
    \includegraphics[width=0.22\textwidth]{theory_3d.eps}
  \label{subfig_theory_3d}
  } 
\end{center}
\end{figure}

\subsection{Experimental Results}
\label{subsec_vs_variables}

In this section, we evaluate the accuracy of the mechanisms defined in
Section (\ref{sec_smoo}) w.r.t. four variables, including data size, dimensions,
data variance, prior distribution, and some combinations thereof.
Every plot is an average over 1000 runs. In all the experiments we set
$\epsilon = 0.8$, and $\delta = 10^{-8}$.

@@

\begin{figure}[H]
\begin{center}
\centering
  \subfigure[Data set size from $300$ to $800$ ]{
    \includegraphics[width=0.45\textwidth]{sampling_2d.eps}
    \label{subsubsec_vs_datasize1a}
  }
\caption{Increasing data size with fixed prior $\betad(1,1)$. Unbalanced datasets of mean $(0.1,0.9)$ and parameters $\epsilon = 0.8$ and $\delta = 10^{-8}$}
\label{fig_vs_datasize}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\centering
  \subfigure[Data set size from $14000$ to $20000$]{
     \includegraphics[width=0.45\textwidth]{sampling_3d.eps}   
  \label{subsubsec_vs_datasize1bdir}
  } 
\caption{Increasing data size with fixed $\dirichlet(1,1,1)$ prior distribution, Unbalanced datasets of mean $(0.2,0.3, 0.5)$ and parameters $\epsilon = 0.8$ and $\delta = 10^{-8}$}
\label{fig_vs_datasize_dir}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\centering
  \subfigure[Data set size from $14000$ to $20000$]{
     \includegraphics[width=0.45\textwidth]{sampling_4d.eps}   
  \label{subsubsec_vs_datasize1bdir}
  } 
\caption{Increasing data size with fixed $\dirichlet(1,1,1)$ prior distribution, Unbalanced datasets of mean $(0.2,0.3, 0.5)$ and parameters $\epsilon = 0.8$ and $\delta = 10^{-8}$}
\label{fig_vs_datasize_dir}
\end{center}
\end{figure}

@@

\paragraph{Fixed dataset varying balanced priors}
\label{subsubsec_vs_prior}
In Figure \ref{fig_vs_prior}, we fix the dataset to be $\langle 5,5,5\rangle$.
We also considered balanced priors with increasing values in their dimensions.
The plot shows that in the beginning the Laplace mechanism performs better but
it is outperformed after a while.
\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{sampling_prior.eps}
\caption{Observed data set is: $(5,5,5)$, varying balanced priors}
\label{fig_vs_prior}
\end{figure}

@@

\begin{figure*}
\begin{center}
\centering
  \subfigure[data size range from 90 to 180]{
    \includegraphics[width=0.21\textwidth]{global_epsilon_20_20.eps}}
  \subfigure[data size range from 90 to 180]{
    \includegraphics[width=0.21\textwidth]{global_epsilon_20_20(2).eps}} 
\caption{Concrete privacy calculation under settings that: prior distribution:$[1,1]$, $\epsilon = 0.8$, $\delta = 0.0005$ and observed data are uniformly distributed}
\label{fig_privacy}
\end{center}
\end{figure*}
